Local simple connectedness of boundaries of hyperbolic groups

### 2020-04-06 paper.tex
Finished proof read

\begin{proposition}
  If $\boundary G - \set\xi$ is path connected then $G$ satisfies $\ddag'_M$
  for any $M$.
\end{proposition}

@@

\begin{proof}
  Let $G$ satisfy $\ddag'_{C+D}$ with constants $L$ and $m_0 \geq n_0$.
  Let $M$ be such that $3\cdot2^M \geq 2L+2$, so that $\sd^M\boundary\Delta$
  has at least $2L+2$ edges.  By increasing $m_0$ we can also assume that $G$
  satisfies $\S_M$ with constants $L'$ and $m_0$.

  Using $\ddag'_{C+D}$ we can extend $\gamma_0\disjointunion\gamma_1$ to a simplicial map
  \begin{equation*}
    J\times\{0\} \union(\Vert J \times \sd^L [0,1]) \union J\times\{1\} \to K_n,
  \end{equation*}
  again thinking of the domain of this map as a subspace of
  $\boundary\Delta\times[0,1]$ in the same way. Each component of the
  complement of this subspace in $\boundary\Delta\times[0,1]$ is a $2$-cell
  whose boundary is an edge path of length $2L+2$. Identifying this edge path
  with a copy of $\sd^M\boundary\Delta$ with some edges collapsed to a point,
  we can use the condition $\S_M$ to extend the map to the full $2$-simplex
  $\Delta$ in such a way that it is affine on $\sd^{L'}\Delta$, and use this
  extension to extend the map to all of $\boundary\Delta\times[0,1]$.

We now need two more lemmas.

@@

@@

\begin{proof}
  By \cref{lem:close_projections} the distances $\newdist(p^\infty_n(\xi),
  p^\infty_{n+1}(\xi))$ are bounded by $6\delta + 3 + 2D$. Let $y$ be a vertex
  of a minimal simplex containing $p^\infty_n(\xi)$; then $\newdist(y,
  p^\infty_n(\xi))$ is bounded by $D$.

@@

\begin{proof}
  Apply Proposition~\ref{prop:basic_homotopies} with $C$ as in
  Lemma~\ref{lem:name_this} to obtain numbers $m_0$ and $\mu$.

@@

  For each $m$ we can apply Proposition~\ref{prop:basic_homotopies} to the maps
  $f_{m+1}$ and $i^{n+m}_{n+m+1}\composed f_m$. To see this, firstly
  Lemma~\ref{lem:name_this} gives the uniformly bounded distance assumption.
  Secondly, $f_{m+1}$ is a simplicial approximation to
  $p^\infty_{n+m+1}\composed\gamma$. Finally,
  $i^{n+m}_{n+m+1}\restricted{\Skel_1 K_{n+m}}$ is a simplicial approximation
  to a map $p^\infty_{n+m+1}\composed j_m$ for some map $j_m\from\Skel_1
  K_{n+m}\to \boundary G$, and then $i^{n+m}_{n+m+1}\composed g_m$ is a
  simplicial approximation to $p^\infty_{n+m+1}\composed j_m\composed f_m$.

@@

  Let $x \in \boundary\Delta$ and let $\epsilon > 0$. There exists $M$ such
  that for $m \geq M$ and $y \in \boundary \Delta$, $\rho(\gamma(y),
  \gamma_m(y)) \leq \epsilon/3$. Increasing $M$ if necessary we may assume that
  $\mu a^{-m-n} \leq \epsilon/3$.
  Finally, let $U$ be a neighbourhood of $x$ such that $\rho(\gamma(x),
  \gamma(y)) \leq \epsilon/3$ for $y \in U$. Now, for any $y \in U$ and 
  $t \in (0, 2^{-M}]$,
  \begin{equation*}
    \rho(h(x,0), h(y,t)) \leq \frac{\epsilon}{3} + \frac{\epsilon}{3} +
        \frac{\epsilon}{3} = \epsilon.
  \end{equation*}
  Therefore $h$ is continuous.

@@

\begin{proof}[Proof of Theorem~\ref{thm:main_theorem}]
  If $G$ is a one-ended hyperbolic group such that $\boundary G - \set\xi$ is
  simply connected for any $\xi \in \boundary G$, then the condition $\S_M$ is
  satisfied for any $M$ by Proposition~\ref{prop:sufficient_condition}. It
  follows that $\boundary G$ is locally simply connected by
  Theorem~\ref{thm:ddagimpliesLCtwo}.

  For the converse, suppose that $\boundary G$ is connected and $\boundary$ 
  is locally simply connected. Let $\xi$ be a point in $\boundary G$ and let
  $\gamma\from S^1\to\boundary G - \set\xi$ be a continuous map.

@@

  Let $U$ be a neighbourhood of $\zeta_+$ such that the closure of $U$ does not
  contain $\zeta_-$ and let $V$ be a neighbourhood of $\zeta_+$ contained in
  $U$ such that any map from the circle to $V$ is nullhomotopic in $U$. Then by
  the properties of the sequence $(g_n)$ described above, there exists an
  element $g_n$ such that $g_n\cdot\xi \notin U$ and $g_n\composed\gamma$ has
  image in $V$. Then $g_n\composed\gamma$ admits a contracting homotopy with
  image in $U$, and the image of this homotopy under $g_n^{-1}$ is a
  contracting homotopy of $\gamma$ with image contaed in $\boundary G -
  \set\xi$.
\end{proof}

@@

\begin{question} Is there a one-ended hyperbolic group $G$ such that $\boundary
  G$ is simply connected but not locally simply connected? Conversely, if
  $\boundary G$ is locally simply connected but not simply connected, does it
  follow that $\boundary G$ is a circle and $G$ is virtually Fuchsian?
\end{question}

>>>>

\begin{proposition}\label{prop:sufficient_for_ddag_p}
  If $\boundary G - \set\xi$ is path connected then $G$ satisfies $\ddag'_M$
  for any $M$.
\end{proposition}

@@

We complete our description of the maps $i^n_\infty$ with the following lemma,
which can be thought of as saying that $i^n_\infty$ is an approximate section
of the map $p^\infty_n$.

@@

\begin{proof}
  Let $G$ satisfy $\ddag'_{C+D}$ with constants $L$ and $m_0 \geq n_0$.
  Let $M$ be such that $3\cdot2^M \geq 2^L+2$, so that $\sd^M\boundary\Delta$
  has at least $2^L+2$ edges.  By increasing $m_0$ we can also assume that $G$
  satisfies $\S_M$ with constants $L'$ and $m_0$.

  Using $\ddag'_{C+D}$ we can extend $\gamma_0\disjointunion\gamma_1$ to a
  simplicial map
  \begin{equation*}
    J\times\{0\} \union(\Skel_0 J \times \sd^L [0,1]) \union J\times\{1\} \to K_n,
  \end{equation*}
  again thinking of the domain of this map as a subspace of
  $\boundary\Delta\times[0,1]$ in the same way. Each component of the
  complement of this subspace in $\boundary\Delta\times[0,1]$ is a $2$-cell
  whose boundary is an edge path of length $2^L+2$. Identifying this edge path
  with a copy of $\sd^M\boundary\Delta$ with some edges collapsed to a point,
  we can use the condition $\S_M$ to extend the map to the full $2$-simplex
  $\Delta$ in such a way that it is affine on $\sd^{L'}\Delta$, and use this
  extension to extend the map to all of $\boundary\Delta\times[0,1]$.

The following proposition extends the family of maps between which we can build
homotopies.

To prove this proposition, we use the following lemma.

@@

\begin{proof}
  By \cref{lem:close_projections} the distances $\newdist(p^\infty_n(\xi),
  p^\infty_{n+1}(\xi))$ are bounded by $6\delta + 3 + 2D$. Let $y$ be a vertex
  of a minimal simplex containing $p^\infty_n(\xi)$.

@@

\begin{proof}[Proof of Proposition~\ref{prop:uniform_homotopies}]
  Apply Proposition~\ref{prop:basic_homotopies} with $C$ as in
  Lemma~\ref{lem:name_this} to obtain numbers $m_0$ and $\mu$.

@@

  For each $m$ we can apply Proposition~\ref{prop:basic_homotopies} to the maps
  $f_{m+1}$ and $i^{n+m}_{n+m+1}\composed f_m$. To see this, firstly
  Lemma~\ref{lem:name_this} gives the uniformly bounded distance assumption.
  Secondly, $f_{m+1}$ is a simplicial approximation to
  $p^\infty_{n+m+1}\composed\gamma$. Finally,
  $i^{n+m}_{n+m+1}\restricted{\Skel_1 K_{n+m}}$ is a simplicial approximation
  to a map $p^\infty_{n+m+1}\composed j_m$ for some map $j_m\from\Skel_1
  K_{n+m}\to \boundary G$, and then $i^{n+m}_{n+m+1}\composed f_m$ is a
  simplicial approximation to $p^\infty_{n+m+1}\composed j_m\composed f_m$.

@@

  Let $x \in \boundary\Delta$ and let $\epsilon > 0$. By
  Lemma~\ref{lem:convergestoidentity} there exists $M$ such that for $m \geq M$
  and $y \in \boundary \Delta$, $\rho(\gamma(y), \gamma_m(y)) \leq \epsilon/3$.
  Increasing $M$ if necessary we may assume that $\mu a^{-M-n} \leq
  \epsilon/3$.  Finally, let $U$ be a neighbourhood of $x$ such that
  $\rho(\gamma(x), \gamma(y)) \leq \epsilon/3$ for $y \in U$. Now, for any $y
  \in U$ and $t \in (0, 2^{-M}]$,
  \begin{equation*}
    \rho(h(x,0), h(y,t)) \leq \frac{\epsilon}{3} + \frac{\epsilon}{3} +
        \frac{\epsilon}{3} = \epsilon.
  \end{equation*}
  Therefore $h$ is continuous.

@@

\begin{proof}[Proof of Theorem~\ref{thm:main_theorem}]
  If $G$ is one-ended then $\ddag'_M$ is satisfied for any $M$ by
  Proposition~\ref{prop:sufficient_for_ddag_p}. Furthermore, if $\boundary G -
  \set\xi$ is simply connected for any $\xi \in \boundary G$ then the condition
  $\S_M$ is satisfied for any $M$ by
  Proposition~\ref{prop:sufficient_condition}. It follows that $\boundary G$ is
  locally simply connected by Theorem~\ref{thm:ddagimpliesLCtwo}.

  For the converse, suppose that $\boundary G$ is connected and locally simply
  connected. Let $\xi$ be a point in $\boundary G$ and let $\gamma\from
  S^1\to\boundary G - \set\xi$ be a continuous map.

@@

  Let $U$ be a neighbourhood of $\zeta_+$ such that the closure of $U$ does not
  contain $\zeta_-$ and let $V$ be a neighbourhood of $\zeta_+$ contained in
  $U$ such that any map from the circle to $V$ is nullhomotopic in $U$. Then by
  the properties of the sequence $(g_n)$ described above, there exists an
  element $g_n$ of the sequence such that $g_n\cdot\xi \notin U$ and
  $g_n\composed\gamma$ has image in $V$. Then $g_n\composed\gamma$ admits a
  contracting homotopy with image in $U$, and the image of this homotopy under
  $g_n^{-1}$ is a contracting homotopy of $\gamma$ with image contaed in
  $\boundary G - \set\xi$.
\end{proof}

@@

\begin{question} Is there a one-ended hyperbolic group $G$ such that $\boundary
  G$ is simply connected but not locally simply connected? Conversely, if
  $\boundary G$ is locally simply connected but not simply connected, does it
  follow that $\boundary G$ is a circle, and so $G$ is virtually Fuchsian?
\end{question}
